---
title: "CSDE 502: Homework #7"
author: "[Esteban J. Valencia](mailto:estebanv@uw.edu)"
date: '`r format(Sys.time(), "%Y-%m-%d %H:%M")`'
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
output: 
    bookdown::html_document2:
        theme: journal
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
urlcolor: blue        
---
```{css, echo=FALSE}
#Step 0.0.1: Setting wide-wise scrolling
pre, code {white-space: pre !important; overflow-x: auto}

```
```{r setup, warning = FALSE, message = FALSE, results = "hide"}
#Step 0.0.2: Specifying chunk options
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      results = "hide",
                      width = 1000)

#Step 0.0.3: Setting CRAN mirror
options(repos = "https://mirror.las.iastate.edu/CRAN/",
        width = 1000)

#Step 0.0.4: Installing & Reading packages
if(!require (knitr)) install.packages("knitr")
library(knitr)
if(!require (bookdown)) install.packages("bookdown")
library(bookdown)
if(!require (tidyverse)) install.packages("tidyverse")
library(tidyverse)
if(!require (magisttr)) install.packages("magrittr")
(library(magrittr))
if(!require (kableExtra)) install.packages("kableExtra")
library(kableExtra)
if(!require (ggplot2)) install.packages("ggplot2")
library(ggplot2)
if(!require (readstata13)) install.packages("readstata13")
library(readstata13)
if(!require (captioner)) install.packages("captioner")
library(captioner)
if(!require (Rmisc)) install.packages("Rmisc")

#Step 0.0.5: Setting captions
cap.tab = captioner(prefix = "Table")
cap.fig = captioner(prefix = "Figure")

#Step 0.0.6: Setting temporary file directory and temporary file
f.dir = paste0(tempdir(), "/")
f.temp = tempfile()

#Step 0.0.7: Downloading/opening Add Health zip-file.

  ##Naming html address
  f.url = "http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip"
  
  ##Setting path for zipfile
  z.file = file.path(Sys.getenv("TEMP"), basename(f.url))
  
  ##Setting file in temporary folder
  f.data = tools::file_path_sans_ext(z.file)
  
  ##Check-loop: Does file exist?
  if(!file.exists(f.data)){
    
    ###If the file does not exist...
    if(!file.exists(z.file)){
      curl::curl_download(url = f.url, destfile = z.file)
    }
    
    ###If the file does exist...
    unzip(ziipfile = z.file, exdir = Sys.getenv("TEMP"))
  }
  
  ##Check-loop: Has file been read? If not, reading in file...
  if(!exists("adh.data")) {
    adh.data = haven::read_dta(f.data)
  }
  
  ##Changing column names to lower case
  colnames(adh.data) %<>% str_to_lower()
```

# Introduction
The following file describes the subsetting and tabulation of data from Add Health. The full Add Health wave 1 data have been downloaded, and consist of `r length(adh.data)` variables. We will parse this down to just 5 variables.

```{r}
#Step 1.0.1: Subsetting data
s.adh = subset(adh.data, 
               select = c(h1ed7, h1ed11, h1ed12, h1ed13, h1ed14))

#Step 1.0.2: Renaming data
s.adh %<>% dplyr::rename(suspensions = h1ed7, 
                         gr_english = h1ed11, 
                         gr_maths = h1ed12, 
                         gr_history = h1ed13, 
                         gr_science = h1ed14)

#Step 1.0.3: Rendering variables as factors
s.adh %<>% lapply(FUN = function(x) factor(x, 
                                           ordered = is.ordered(x),
                                           labels = attributes(unlist(attributes(x)$labels))$names))
s.adh = data.frame(s.adh)
```

Once the data frame is subset, we can examine the variables we have selected by variable name and variable label. 

```{r}
#Step 1.0.4: Specifying metadata
m.data = bind_cols(varname = colnames(s.adh),
                   varlabel = s.adh %>% map(~attributes(.)$label) %>% unlist())

#Step 1.0.5: Printing variable table
DT::datatable(m.data)

```

# Tables

Below presents a frequency tabulation of the variable `$suspensions`.

```{r, results = "markdown"}
#Step 2.0.1: Tabulation
tab.x = table(s.adh$suspensions)
tab.y = round(prop.table(tab.x) * 100, digits = 1)
tab.z = cbind(N = tab.x, Percent = tab.y)
kable(tab.z) %>%
  kable_styling(full_width = FALSE,  position = "left")
```

Below is a frequency tabulation of the variable `$gr_english`.

```{r, results = "markdown"}
#Step 2.0.2: Tabulation
tab.x = table(s.adh$gr_english)
tab.y = round(prop.table(tab.x) * 100, digits = 1)
tab.z = cbind(N = tab.x, Percent = tab.y)
kable(tab.z) %>%
  kable_styling(full_width = FALSE,  position = "left")
```

Below is a frequency tabulation of the variable `$gr_maths`.

```{r, results = "markdown"}
#Step 2.0.3: Tabulation
tab.x = table(s.adh$gr_maths)
tab.y = round(prop.table(tab.x) * 100, digits = 1)
tab.z = cbind(N = tab.x, Percent = tab.y)
kable(tab.z) %>%
  kable_styling(full_width = FALSE,  position = "left")
```

Below is a frequency tabulation of the variable `$gr_history`.

```{r, results = "markdown"}
#Step 2.0.4: Tabulation
tab.x = table(s.adh$gr_history)
tab.y = round(prop.table(tab.x) * 100, digits = 1)
tab.z = cbind(N = tab.x, Percent = tab.y)
kable(tab.z) %>%
  kable_styling(full_width = FALSE,  position = "left")
```

Below is a frequency tabulation of the variable `$gr_science`.

```{r, results = "markdown"}
#Step 2.0.5: Tabulation
tab.x = table(s.adh$gr_science)
tab.y = round(prop.table(tab.x) * 100, digits = 1)
tab.z = cbind(N = tab.x, Percent = tab.y)
kable(tab.z) %>%
  kable_styling(full_width = FALSE,  position = "left")
```

Below is a frequency cross-tabulation of the variables `$suspensions` & `$gr_english`.

```{r, results = "markdown"}
#Step 2.0.6: Tabulation

  ##Tabulation for No Suspensions
  tab.x1 = table(s.adh$gr_english[s.adh$suspensions == "(0) No (skip to Q.9)"])
  tab.y1 = round(prop.table(tab.x1) * 100, digits = 1)
  tab.z1 = cbind(N = tab.x1, Percent = tab.y1)
  rownames(tab.z1) = paste0(rownames(tab.z1), ":  Never Suspended")
  
  ##Tabulation for Yes Suspensions
  tab.x2 = table(s.adh$gr_english[s.adh$suspensions == "(1) Yes"])
  tab.y2 = round(prop.table(tab.x2) * 100, digits = 1)
  tab.z2 = cbind(N = tab.x2, Percent = tab.y2)
  rownames(tab.z2) = paste0(rownames(tab.z2), ":  Previously Suspended")
  
  ##Adding tabulations together
  tab.fin = rbind(tab.z1, tab.z2)
  
  ##Printing table
  kable(tab.fin) %>%
  kable_styling(full_width = FALSE,  position = "left")
```

# Plot

The below barplot graphs the distribution of `$suspesions`

```{r, results = "markdown"}
#Step 3.0.1: Recoding suspension variable
attributes(s.adh$suspensions)$levels = c("(0) Never Suspended",
                                         "(1) Prior Suspension",
                                         "(6) Refused to Answer",
                                         "(8) Don't Know Answer")

#Step 3.0.2: Plotting distribution
ggplot(data = s.adh, mapping = aes(x = suspensions)) +
  geom_bar(aes(y = (..prop..) * 100, group = 1)) +
  ylab("Response Proportion (%) \n") +
  xlab("\nHistory of Suspension, Survey Response \n") +
  labs(title = "Figure 1. Distribution of Responses to Suspension Question, Add Health Data")
```

# Saving data

The data are saved as an `RDS` object in a temporary directory.

```{r}
#Step 4.0.1: Saving data
saveRDS(s.adh, paste0(f.dir, "adh_data.rds"))
```

# Source code

The source code for this document is indicated below

```{r, results = "markdown"}
cat(readLines(con = "estebanv_csde_502_hw7.Rmd"), sep = '\n')
```